Score-based tests of measurement invariance: Use in practice

### 2014-07-08 frontiers/IMPSpres-knitr.Rnw
update presentation according to Achim's comments.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@352 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{beamer}
\setbeamertemplate{caption}[numbered]
\usepackage{beamerthemeSzeged,graphics,amsmath,bm,booktabs}
\usepackage{appendixnumberbeamer}
\usepackage[notocbib]{apacite}
\usepackage{tikz}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\setbeamercolor{block title}{use=structure,fg=black,bg=blue!10!white}
\setbeamercolor{block body}{use=structure,fg=black,bg=blue!05!white}
\usefonttheme{serif}
% Seems necessary for texlive 2012?
\usetikzlibrary{arrows}

@@

\section{Theoretical Background}
\begin{frame}
  \frametitle{Score-based Tests}
 \begin{itemize}
    \item The score-based tests \cite{MerZei13, MerFanZei}
      rely on first derivatives of
      the model's log-likelihood function.
    \item We consider individual terms ({\em scores}) of the gradient.  
      These scores tell us how well a particular parameter describes a 
      particular individual.
  % Equation of gradient vs scores for individuals
  \end{itemize}

@@

\begin{frame}
  \frametitle{Visual Explanation of the Score-based Test}
  \begin{columns}
    \begin{column}{0.48\textwidth}
      \begin{block}{Score}
        \resizebox{5.0cm}{!}{
        $$\bordermatrix{\text{}& \theta_{1} & \theta_{2} & \ldots 
          &\textcolor{red}{\theta_{k}}\cr
                x_{1}&  1.32  & -4.32 & \ldots & -2.34\cr
                x_{2}&  3.37 & 8.32 & \ldots & 63.45\cr
                x_{3}&  4.12 & 3.48 & \ldots & -3.11\cr
                x_{..}& \vdots & \vdots & \vdots & \vdots\cr
                x_{n-1}&  12.34 & 13.32 & \ldots & 6.89\cr
                x_{n}& -5.78     &   3.50     & \ldots & 7.67} $$}
        \end{block}
     \end{column}
     \begin{column}{0.48\textwidth}
       \begin{block}{Empirical Cumulative Score}
         \resizebox{5.0cm}{!}{
      $$\bordermatrix{\text{}& \theta_{1} & \theta_{2} & \ldots
        &\textcolor{red}{\theta_{k}}\cr
                x_{1}&  1.32  & -4.32 & \ldots & -2.34\cr
                x_{2}&  4.69 & 4.00 & \ldots & 61.11\cr
                x_{3}&  8.81 & 7.48 & \ldots & 58.00\cr
                x_{..}& \vdots & \vdots & \vdots & \vdots\cr
                x_{n-1}&  7.89 & 3.32 & \ldots & 4.91\cr
                x_{n}& 0     &   0     & \ldots & 0} $$}
        \end{block}
      \end{column}
   \end{columns}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[1.]
 \item Input data
\end{enumerate}
\begin{itemize}
  \item Gratitude scale (GQ-6) completed by six age
    groups(10-11, 12-13, 14, 15, 16, 17-19) \cite{frofan11}
  \item $n=1327$.
  \item LRT is always significant when sample size is large. 
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[2.]
  \suite
\item Fit the constrained model: no need to compare various models.
\end{enumerate}
<<modeest,eval=FALSE>>=
 library(lavaan)
  restr <- cfa("f1 =~ gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
               data = yg, group = "agegroup", 
               meanstructure = TRUE, group.equal = "loadings")
@

@@

\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[3.]
  \suite
  \item Test of measurement invariance
\end{enumerate}
<<sctest-ord2, eval=FALSE>>=
 library(strucchange)
maxlmo <- sctest(restr, order.by = yg$agegroup, parm = 1:4, 
                 vcov = "info", functional = "maxLMo",
                 plot = TRUE)
@
<<sctest-ord2res, eval=TRUE>>=
maxlmo
@ 
\end{frame}

@@

\begin{frame}
\begin{figure}
\tiny{\caption{Fluctuation processes for the $\mathit{LM}_o$ statistic}}
\label{fig:ordres}
<<sctest-ord-graphics, fig.height=3.3, fig.width=5, echo=FALSE, results='hide'>>=
sctest(restr, order.by = yg$agegroup, parm = 1:4, vcov = "info",
  functional = maxLMo, plot = TRUE, 
               ylim = c(0, 14), xlab = "Age group", 
               ylab = "LM statistics")
@
\end{figure}
\end{frame}

>>>>

\documentclass{beamer}
\setbeamertemplate{caption}[numbered]
\usepackage{beamerthemeSzeged,graphics,amsmath,bm,booktabs}
\usepackage{appendixnumberbeamer}
\usepackage{blkarray}
\usepackage[notocbib]{apacite}
\usepackage{tikz}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\setbeamercolor{block title}{use=structure,fg=black,bg=blue!10!white}
\setbeamercolor{block body}{use=structure,fg=black,bg=blue!05!white}
\usefonttheme{serif}
% Seems necessary for texlive 2012?
\usetikzlibrary{arrows}

@@

\section{Theoretical Background}
\begin{frame}
  \frametitle{Score-based Tests}
 \begin{itemize}
  %  \item The score-based tests \cite{MerZei13, MerFanZei}
  %    rely on first derivatives of
  %    the model's log-likelihood function.
    \item Access deviations in the model score \cite{MerZei13,
      MerFanZei}, i.e., the first derivatives of the model's log
      likelihoood function.  
    \item We consider individual terms ({\em scores}) of the gradient.  
      These scores tell us how well a particular parameter describes a 
      particular individual.
  % Equation of gradient vs scores for individuals
  \end{itemize}

@@

\begin{frame}
  \frametitle{Visual Explanation of the Score-based Test}
  \begin{columns}
    \begin{column}{0.48\textwidth}     
      \begin{block}{Score}
        \resizebox{5.0cm}{!}{ 
          $
            \begin{blockarray}{cccccc}
              \text{}& \theta_{1} & \theta_{2} & \ldots 
                     &\textcolor{red}{\theta_{k}} & \text{age} \\
               \begin{block}{l(cccc)c}
                 x_{1} &  1.32  & -4.32 & \ldots & -2.34 & 10.5\\
                 x_{2} &  3.37 & 8.32 & \ldots & 63.45 & 11.0\\ 
                 x_{3} &  4.12 & 3.48 & \ldots & -3.11 & 11.3\\
                 x_{..} & \vdots & \vdots & \vdots & \vdots & 17.1\\ 
                 x_{n-1} &  12.34 & 13.32 & \ldots & 6.89 & 18.6\\
                 x_{n} & -5.78 & 3.50 & \ldots & 7.67 & 18.9\\
               \end{block}
            \end{blockarray} 
           $
          }
        \end{block}   
     \end{column}

     \begin{column}{0.48\textwidth}
          \begin{block}{Empirical Cumulative Score}
          \resizebox{5.0cm}{!}{
           $
            \begin{blockarray}{cccccc}
              \text{}& \theta_{1} & \theta_{2} & \ldots 
                     &\textcolor{red}{\theta_{k}} & \text{age} \\
               \begin{block}{l(cccc)c}
                 x_{1} &  1.32  & -4.32 & \ldots & -2.34 & 10.5 \\
                 x_{2} &  4.69 & 4.00 & \ldots & 61.11 & 11.0 \\
                 x_{3} &  8.81 & 7.48 & \ldots & 58.00 & 11.3 \\
                 x_{..} & \vdots & \vdots & \vdots & \vdots & 17.1 \\
                 x_{n-1} &  7.89 & 3.32 & \ldots & 4.91 & 18.6 \\
                 x_{n}& 0     &   0     & \ldots & 0 & 18.9 \\
               \end{block}
            \end{blockarray} 
           $
         }  
        \end{block}
      \end{column}
   \end{columns}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[1.]
 \item Input data
\end{enumerate}
\begin{itemize}
  \item Gratitude scale (GQ-6) completed by six age
    groups (10--11, 12--13, 14, 15, 16, 17--19) \cite{frofan11}
  \item $n=1327$.
  \item LRT is always significant when sample size is large. 
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[2.]
  \suite
\item Fit the constrained model: no need to compare various models.
\end{enumerate}
<<modeest,eval=FALSE>>=
 library("lavaan")
  restr <- cfa("f1 =~ gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
               data = yg, group = "agegroup", 
               meanstructure = TRUE, group.equal = "loadings")
@

@@

\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[3.]
  \suite
  \item Test of measurement invariance
\end{enumerate}
<<sctest-ord2, eval=FALSE>>=
 library("strucchange")
maxlmo <- sctest(restr, order.by = yg$agegroup, parm = 1:4, 
                 vcov = "info", functional = "maxLMo",
                 plot = TRUE)
@
<<sctest-ord2res, eval=TRUE>>=
maxlmo
@ 
\end{frame}

@@

\begin{frame}
\begin{figure}
\tiny{\caption{Fluctuation processes for $\max\mathit{LM}_o$ statistic}}
\label{fig:ordres}
<<sctest-ord-graphics, fig.height=3.3, fig.width=5, echo=FALSE, results='hide'>>=
par(cex = 0.8)
sctest(restr, order.by = yg$agegroup, parm = 1:4, vcov = "info",
  functional = maxLMo, plot = TRUE,  
               ylim = c(0, 14), xlab = "Age group", 
               ylab = "LM statistics", main = expression(maxLM[o]))
@
\end{figure}
\end{frame}
